---
title: "Using skygrowth on a simulated tree"
author: "Xavier Didelot"
date: '`r Sys.Date()`'
output:
  pdf_document: default
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Using skygrowth on a real dataset}
  %\usepackage[utf8]{inputenc}
---

# Initialisation

In this vignette we demonstrate the usage of `skygrowth` on coalescent simulated trees.
We need to load the packages `skygrowth` and `ape`, and we set the random seed to 0 to make the vignette reproducible. 

```{r}
library(skygrowth)
library(ape)
set.seed(0)
```

# Simulation

Let us consider a pathogen population with effective population size times 
generation time equal to $N_e g=10$ years.
A sample of 100 individuals is taken from this population at a single timepoint.
The genealogy of such a sample can be simulated as follows:


```{r}
tree=rcoal(100)
tree$edge.length=tree$edge.length*10
plot(tree, show.tip.label = F)
axisPhylo(backward = T)
mtext("Years since most recent sample", side=1, line=3)
```

# MAP analysis

```{r}
fit=skygrowth.map(tree,quiet=T)
cat('Drift parameter tau: ',fit$tau)
growth.plot(fit)
plot(fit,logy=F)
```

# MCMC analysis with Gibbs move

```{r}
fit2=skygrowth.mcmc(tree,quiet=T)
growth.plot(fit2)
plot(fit2,logy=F)
plot(fit2$tau,type='l',ylab='tau',xlab='Iterations')
```

# MCMC analysis with MH move

```{r}
fit3=skygrowth.mcmc(tree,quiet=T,tau_logprior = function(x) dexp(x, 0.1, log=T))
growth.plot(fit3)
plot(fit3,logy=F)
plot(fit3$tau,type='l',ylab='tau',xlab='Iterations')
```

# mlesky analysis

```{r}
library(mlesky)
fit4=mlskygrid(tree)
plot(fit4,logy=F)
```

# phylodyn analysis

```{r}
library(phylodyn)
fit5=BNPR(tree)
plot_BNPR(fit5)
```